---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
df <- read.csv("G:/RProg/activity.csv")
dfNA <- df[!is.na(df$steps), ]
```


## What is mean total number of steps taken per day?

```{r, echo=FALSE}
library(plyr)
stepsByDay <- ddply(dfNA, .(date), summarise, steps=sum(steps))
hist(stepsByDay$steps, main="Number of Steps", xlab="Steps taken per day")
```

####Mean and Median of the total number of steps taken per day

```{r}
mean(stepsByDay$steps)
median(stepsByDay$steps)
```

## What is the average daily activity pattern?

```{r, echo=FALSE}
averageByInterval <- ddply(dfNA, .(interval), summarise, steps = mean(steps))
plot(averageByInterval$interval, averageByInterval$steps, type = "l", xlab = "5-minute Interval", ylab = "Average Number of steps taken", main = "Daily activity pattern")
```

#### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
averageByInterval[averageByInterval$steps == max(averageByInterval$steps), ]
colnames(averageByInterval)[2] <- "intervalAvg"
```


## Imputing missing values

#####Total number of rows with NAs
```{r}
nrow(df[is.na(df$steps), ])
```

#####Fill NAs with mean for that 5-min interval
```{r, results='hide'}
dfNoNA <- arrange(join(df, averageByInterval), interval)
dfNoNA$steps[is.na(dfNoNA$steps)] <- dfNoNA$intervalAvg[is.na(dfNoNA$steps)]
```


#####Recalculate the average daily pattern
```{r, echo=FALSE}
stepsByDayNoNA <- ddply(dfNoNA, .(interval), summarise, steps = mean(steps))
hist(stepsByDayNoNA$steps, main="Number of Steps", xlab="Steps taken per day")
```

####Mean and Median of the total number of steps taken per day

```{r}
mean(stepsByDayNoNA$steps)
median(stepsByDayNoNA$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lattice)
weekdays <- weekdays(as.Date(dfNoNA$date))
dfWeekdays <- transform(dfNoNA, day = weekdays)
dfWeekdays$wk <- ifelse(dfWeekdays$day %in% c("Saturday", "Sunday"), 
    "weekend", "weekday")
averageIntervalByweek <- ddply(dfWeekdays, .(interval, wk), summarise, 
    steps = mean(steps))

xyplot(steps ~ interval | wk, data = averageIntervalByweek, layout = c(1, 2), 
    type = "l")
```